### 2024-05-31 manuscript/manuscript.qmd
cleaning up manuscript folder

Previous research has shown that people who grew up in adverse conditions tend to be slower on the Flanker Task, suggesting difficulty inhibiting distractions.
However, this pattern might result from slower general cognitive processes (e.g., reduced general processing speed), rather than the specific ability of inhibition.
We used Drift Diffusion Modeling in three studies---one pilot and two confirmatory studies---to better understand associations of adversity exposures with Flanker performance. 
Neither exposure to violence nor unpredictability were associated with inhibition.
In contrast, although mixed, exposure to violence and unpredictability were associated with lower strength of perceptual input; that is, how well someone can process the arrows in general.
Finally, lower strength of perceptual input was linked to a tendency to process information more holistically, focusing less on individual details.
Thus, researchers should be cautious when drawing inferences about inhibition abilities in people exposed to adversity based on response times alone.

@@

The overarching goal of our studies is to understand the attentional and processing styles that people develop in conditions of adversity.
We focus on measures of exposure to violence and environmental unpredictability.
Previous research shows that these two types of adversity are on the one hand associated with improved attention shifting and working memory updating [@fields_2021; @mittal_2015; @young_2018; @young_2022], and on the other hand with lowered inhibition and working memory capacity [@fields_2021; @mittal_2015; @young_2018].
We conducted three online studies: one pilot and two follow-up studies.
Using cognitive modeling, we unpack Flanker performance in comparison to other tasks that require externally focused attention (Pilot), across visual processing manipulations (Study 1), and in terms of tendencies for holistic versus local processing (Study 2).

We used an incremental preregistration approach across studies (for all preregistrations, code and materials, see [ANONYMIZED FOR REVIEW].
For each study, we preregistered primary (i.e., hypothesis-driven) and exploratory analyses.
The main text addresses the primary analyses involving violence exposure and the exploratory analyses involving environmental unpredictability.
We describe the other exploratory analyses in the supplemental materials (section 2).
For an overview of all deviations from the preregistrations, see section 4 of the supplemental materials.

# Pilot

In the Pilot, our goal was to understand how childhood adversity relates to performance on tasks with different attentional demands.
Participants completed self-report measures of childhood adversity and three cognitive tasks.
These tasks measured inhibition, attention for peripheral cues, and attention for subtle changes.
We expected violence exposure to be positively associated with a present-oriented attention style.
Specifically, we expected participants with more violence exposure to be better at detecting peripheral stimuli and subtle changes.
We expected this would result in a higher drift rate (faster speed of information accumulation) or shorter non-decision times (faster attention orientation, among other things), but not necessarily with differences in boundary separation (response caution).
Second, we expected that participants with more violence exposure would be worse at ignoring distracting peripheral stimuli.
We expected this would result in more experienced interference (as derived from the SSP model).

@@

Participants were 565 US-based individuals aged between 18 and 30 recruited on Prolific Academic ([https://www.prolific.co](https://www.prolific.co)) (See Table 1 for demographic data).
The sample was balanced on sex.
We used the MacArthur's ladder, included in Prolific's prescreening battery, for assessing perceived SES to ensure about half of the sample came from lower-SES backgrounds (which we defined as a score of 4 or below).
Participants were eligible if they spoke fluent English and did not report color-blindness.

@@

We conducted a power simulation using the *faux* package in R [@debruine_2021] to determine the minimally required number of participants for standardized regression coefficients of 0.10 and 0.15 (for details and simulation code, see [ANONYMIZED FOR REVIEW]).
Power was \> .80 for adversity x task condition interactions with *N* = 450 or more.
For a linear main effect, detecting an effect of $\beta$ = 0.15 with .90 power would require *N* = `r pwr.r.test(r = 0.15, sig.level = 0.05, power = .90, alternative = "two.sided")[['n']] %>% round(0)`.
We sampled 550 participants, anticipating a final sample of \~500 after exclusions.

Prior to analyzing the data, we applied our preregistered exclusion criteria.
First, we excluded participants who did not complete the full study; second, those who had incomplete data on any of the attention tasks; third, those who missed both attention check items; fourth, those who had suspicious response patterns (e.g., consistently endorsing high response options even though some items were reverse coded).
Fifth, on a trial-level, we excluded any trials with reaction times \< 250 ms or \> 3500 ms [@ratcliff_2015].
Participants with more than 10 trials removed were completely excluded from the analyses.\
The final sample consisted of `r nrow(pilot_data)` participants.

@@

The pattern of findings on the Flanker was interesting for two reasons.
First, the Flanker is a widely used task to assess the ability to inhibit irrelevant information, and people exposed to adversity typically show lowered performance.
Our pilot results, though, suggest that lowered performance may not be caused by a reduced ability to inhibit distracting information.
Instead, people exposed to adversity might have a lower strength of perceptual input, leading to slower and less efficient information processing.
If true, these initial findings suggest that performance might be improved through interventions that increase the visual quality of stimuli.
In Study 1, we aimed to replicate and extend these findings.

@@

The goal of Study 1 was to follow up on the Pilot by manipulating the visual quality of information on the Flanker.
Participants completed three versions: a standard version (similar to the Pilot), one with enhanced visual information, and one with degraded visual information.
We again focused on childhood exposure to violence.
Our first aim was to examine the robustness of our finding of improved interference control on the Flanker in relation to more adversity exposure in the pilot.
We did so by analyzing the data of the standard condition, as well as by pooling the data of the Pilot and Study 1.
Our second aim was to investigate whether manipulating visual information on the Flanker would influence performance for people with more violence exposure.

We preregistered two potential data patterns and associated interpretations, without favoring one over the other *a priori*.
First, the strength of perceptual input might be lower for people with more exposure to violence compared to people with less exposure to violence across all conditions.
Second, lower performance on the standard version might reflect an adaptive trade-off towards cognitive functioning that is less affected by noise or perturbations, at a cost of lower overall performance [@giudice_2018].
In that case, we would expect the strength of perceptual input to be influenced to a lesser extent across conditions for people with more exposure to violence than for people with less exposure to violence.
As a result, they might not benefit as much from enhanced visual information, yet might be able to better maintain performance with degraded information.

@@

Participant recruitment was identical to the Pilot.
In total, 567 people from the United States between the ages of 18 and 30 participated (See Table 1).
We applied the same exclusion criteria as reported in the Pilot.
The final sample consisted of `r nrow(study1_data)` participants.

### Flanker Task

We programmed the Flanker in JsPsych version 6.3.1 [@deleeuw_2015] with three conditions.
Each condition consisted of eight practice trials, followed by 64 test trials.
In the *standard* condition, the arrows were 40 pixels in size (0.4 inches) and had zero padding between them.
In the *enhanced* condition, we increased the arrow size by 12.5% to 45 pixels (0.45 inches), and increased the space between the arrows to 5 pixels.
This increased the width of the stimulus display by 50% with respect to the standard display.
In the 'degraded' condition, sizes and space between arrows were the same as in the standard version, but all arrows were rotated 45$^\circ$ (in the same direction; that is, if the left flanker arrows pointed upwards, the right flanker arrows pointed downward).
Participants completed each condition separately in different blocks, in randomized order.

@@

The self-report measures were identical to those used in the Pilot.

@@

The procedure was identical to the pilot study.
The full experiment took approximately 30 minutes.
Participants were paid £3.75 after they completed the full study.

@@

**Multiverse analysis.** We included the same arbitrary decisions in the multiverse analyses as in the Pilot.
For the pooled analyses---i.e., joint analysis of the Pilot and the standard condition of Study 1---there was one minor change in how we included screen rescaling as a preprocessing decision in the multiverse.
In Study 1, we changed the screen rescaling procedure by converting the initial size of the resize box to 300 pixels instead of 100 pixels.
This way, the stimulus display would still be close to the intended size if participants did not engage in any resizing.
However, this led to one important change for the pooled analysis: rescaling (yes or no) was included as an arbitrary exclusion decision in the multiverse analyses with four combinations: (1) exclude non-scalers in both studies; (2) include non-scalers in both studies; (3) exclude non-scalers in the Pilot, include non-scalers in Study 1; (4) include non-scalers in the Pilot, exclude non-scalers in Study 1.

For each analysis, we report the median $\beta$s, 95% confidence intervals, and the proportion of *p*-values \< .05 across all analytic decisions.
For the primary analyses, we used bootstrapping to compute the probability of obtaining an effect size at least as extreme as observed in the real data, conditioned on a true effect size of zero [for details, see @simonsohn_2020].
See the supplemental materials (section 5) for figures summarizing *p*-distributions and the explained variance in the regression coefficients of each data cleaning decision.

**Primary analyses.** To address the first aim, we analyzed the data from the standard condition, as well as pooled the Flanker data of the Pilot and the current study.
We ran separate linear models for each SSP parameter as well as RT difference scores with violence exposure as main predictor and study as covariate (effect-coded).
To address the second aim, we analyzed the effect of violence exposure and Flanker condition type on performance using linear mixed effects models with a random intercept per participant.
The five main dependent variables were mean RT difference and the SSP parameters: Perceptual input, boundary separation, non-decision time, and interference.
For each outcome measure, we ran two separate models: one comparing the standard condition with the enhanced condition, and one comparing the standard condition with the degraded condition.
In both models, condition was dummy-coded using the standard condition as the reference group.

@@

Table 4 summarizes the multiverse results for the effects of violence exposure (primary analysis) and unpredictability (exploratory analysis).
Unlike in the Pilot, we did not find any significant associations with violence exposure.
In the exploratory analysis, there was a significant negative association between unpredictability and perceptual input (median $\beta$ = `r study1_expl_results_list$p_flanker$median_effect_Main`, 95% CI = `r study1_expl_results_list$p_flanker$CI`, `r study1_expl_results_list$p_flanker$p_sum` % of *p*s \<.05).

@@

After pooling the data of the pilot study and Study 1, there was a negative association between violence exposure and interference (median $\beta$ = `r study1_pooled_results_list$interference_vio$median_effect`, 95% CI = `r study1_pooled_results_list$interference_vio$CI`, `r study1_pooled_results_list$interference_vio$p_sum` % of *p*s \<.05, bootstrapped *p* = `r study1_pooled_results_list$interference_vio$boot_p`).
Violence exposure was associated with lower strength of perceptual input under `r study1_pooled_results_list$interference_vio$p_sum`% of multiverse specifications, but the bootstrapped *p*-value was not significant (median $\beta$ = `r study1_pooled_results_list$p_vio$median_effect`, 95% CI = `r study1_pooled_results_list$p_vio$CI`, bootstrapped *p* = `r study1_pooled_results_list$p_vio$boot_p`).
We did not find other significant associations for either violence exposure or unpredictability.

### Flanker conditions

@@

The results of the Pilot and Study 1 were inconsistent with regard to the association between adversity and interference, but hinted at two general patterns.
First, violence exposure was not associated with *increased* interference; instead, we found either the opposite effect or no effect.
Second, both violence exposure and unpredictability were associated with lowered strength of perceptual input, albeit inconsistently.
These findings, if replicable, are intriguing as they would suggest that the common finding of lowered Flanker performance among people with more adversity exposure do not actually result from worse interference control---as typically inferred.
Rather, such lowered performance would result from processes other than inhibition ability, such as slower general processing.
Though interesting, our findings so far leave open the question why adversity might be negatively associated with perceptual input.
This question was the focus of Study 2.

@@

Study 2 set out to compare two explanations for the finding that people exposed to adversity tended to show lower strength of perceptual input on the Flanker.
First, lowered strength of perceptual input in people exposed to adversity may indicate a difficulty in extracting relevant information (i.e., about their direction) from the arrows in general.
Second, the difference in perceptual input may not be a cognitive deficit per se, but instead could be a signature of a difference in processing style---that is, a feature, and not a bug.
People exposed to adversity may process information more holistically, focusing more on the configuration of pieces of information rather than individual pieces of information.
On the Flanker, this would lower the depth of perceptual processing of any individual stimulus, thus resulting in lowered strength of perceptual input, as we observed in the Pilot and Study 1.

We preregistered three aims focusing both on violence exposure and unpredictability.
First, we expected to replicate our earlier findings that adversity was associated with lowered perceptual input and lower interference in people exposed to adversity.
Second, we included a Global-Local Task to investigate the hypothesis---based on the findings of the Pilot and Study 1---that people with more adverse experiences would develop a more holistic style of information processing.
Third, we planned to conduct a within-subjects analysis of Flanker and Global-Local performance to assess whether people with lowered perceptual input on the Flanker Task would also show a more global processing style (rather than a local processing style) on the Global-Local Task.

@@

Participants were 600 people from the United States between the ages of 18 and 30.
Recruitment was identical to Study 1.
We conducted a simulation-based power analysis for the planned linear mixed models with the Global Local Task (see GitHub).
We determined that power of \> .80 for a standardized interaction effect of 0.06, with sigma (noise) set to 0.7 (comparable to observed sigmas in the first two studies) would require 550 participants.
We recruited 600 participants, with the expectation to have a final sample of 550 participants after exclusions.
We applied the same exclusion criteria as reported in the Pilot and Study 1.
The final sample consisted of `r nrow(study2_data)` participants.

@@

The measures of childhood violence exposure and environmental unpredictability were identical to Study 1.
The Flanker was identical to the standard version used in Study 1.

*Global-Local Task*.
The Global-Local Task is a measure of global-local processing [@navon_1977].
Many different versions of this task exist in the literature.
One key dimension on which they differ is whether the task measures focused attention (by cueing attention towards the global or local level prior to stimulus presentation) or divided attention (by having participants search for a target on both levels) [@lee_2023].
Here, we use a version measuring divided attention, which allows measuring whether someone tends to have a more global versus local processing style [@hakim_2017; @lee_2023; @mckone_2010].

Participants saw images of big, black letters (the global level) comprised of small letters (the local level)---so-called Navon images [@navon_1977]---against a white background.
Participants first completed eight practice trials, after which they completed an additional 64 test trials.
On each trial, participants searched for one of two target letters---an 'E' or 'H'---and indicated whether it was present on the global or local letter by pressing 'g' or 'l' on their keyboard, respectively.
Each stimulus was 600 pixels high and 395 pixels wide and comprised of seven local letters vertically and five local letters horizontally.
All stimuli consisted of combinations of the letters 'T', 'F', 'P', 'L', 'H', and 'E'.
The target letters 'H' and 'E' were both present in each stimulus, but the other letters varied.
The global and local level never contained the same letter.

@@

**Multiverse analysis.** We included the same decisions in the multiverse as in the previous studies.
However, there was one deviation from the preregistration: the multiverse analysis contained the same arbitrary decisions as the Pilot and Study 1, instead of a subset, as we preregistered (for details, [ANONYMIZED FOR REVIEW]).
See the supplemental materials (section 5) for figures summarizing *p*-distributions and the explained variance in the regression coefficients of each data cleaning decision.

**DDM estimation.** For the Flanker, we used the SSP [@grange_2016; @white_2018a; @white_2018b; @white_2011] using the same procedure as in Study 1. For the Global-Local Task, we used a hierarchical Bayesian DDM to fit the data using the *runjags* package [@denwood_2016]. See the supplemental materials (Section 3) for more information about the procedure and model fit.

We deviated from our preregistration regarding the preprocessing of Global-Local Task data.
Specifically, we relaxed the low performance threshold as the task was more difficult than anticipated.
These deviations are described in the supplemental materials (section 4).

**Primary analyses.** We ran simple regressions for analyses involving only main effects (aim 1), and linear mixed effects models for analyses involving within-subject interactions (aim 2 and 3).
To address aim 3 (within-subject interaction between Global-Local drift rate and Flanker strength of perceptual input), we further preprocessed the data in two steps.
First, we computed a difference score of Global-Local drift rates by subtracting the drift rate on local trials from the drift rate on global trials (with higher scores reflecting relatively faster information processing on global trials).
Second, we separately standardized the Flanker strength of perceptual input and Global-Local drift rate difference.
We fitted linear mixed effects models with the standardized performance measures as the dependent variable, and adversity type, task (Flanker or Global-Local, sum-coded) and their interaction as independent variables.

@@

```{r}
#| label: figure3
#| fig.width: 7
#| fig.height: 9
#| dpi: 600
#| fig-cap: | 
#|   **Figure 3.** Multiverse results for the association between unpredictability with the strength of perceptual input and interference on the Flanker across all studies. Each panel depicts sorted beta coefficients across all combinations of arbitrary decisions (i.e., the effect curve across the whole multiverse). The top row depicts effect curves in the Pilot. The second row depicts effect curves in Study 1. The third row depicts effect curves in Study 2. The Fourth row depicts effect curves of the pooled analyses across all studies.
knitr::include_graphics("figures/fig3.png")
```

@@

### Global-Local Task performance

@@

To sum up, Study 2 provided additional support for the basic finding that violence exposure and unpredictability were associated with lower strength of perceptual input but not with differences in interference; with the caveat that the associations for unpredictability only showed up in pooled analyses.
People with more exposure to violence and unpredictability also processed information more slowly on the Global-Local Task.
In line with our expectations, childhood exposure to violence was associated with both lowered strength of perceptual input and a more holistic processing style.
The same processing style was observed for participants with more exposure to unpredictability, although they did not show lowered strength of perceptual input.

@@

**Drift Diffusion Model:** A formal cognitive model that estimates cognitive processes at different stages of the decision-making process for speeded binary decision-making tasks.
It represents decision-making as a process in which people accumulate information until one response is sufficiently favored over the other.

@@

**Perceptual input:** Parameter from the Shrinking Spotlight Model of Flanker performance. Measures how well a person can process the Flanker stimuli in general.

@@

**Shrinking Spotlight Model:** An extension of the standard Drift Diffusion Model to account for attention processes on the Flanker.

>>>>

It is well-established that individuals who grew up in adverse conditions tend to be slower on the Flanker Task. 
This finding is typically interpreted to reflect difficulty inhibiting distractions. 
However, it might result from slower general cognitive processes (e.g., reduced general processing speed), rather than the specific ability of inhibition. 
We used Drift Diffusion Modeling in three online studies (total N = 1560) with young adults to understand associations of adversity with Flanker performance. 
We find no associations between exposure to violence and unpredictability with inhibition. 
Yet, although mixed, violence and unpredictability exposure were associated with lower strength of perceptual input--—how well someone can process target and distractor information alike. 
Finally, people with lower strength of perceptual input processed information more holistically, focusing less on details. 
Thus, lowered Flanker performance does not necessarily imply lowered inhibition ability. 
Cognitive modeling reveals a different picture of abilities in adverse conditions as opposed to analyses based on raw performance.

@@

The overarching goal of our studies is to understand the attentional and processing styles that people develop in conditions of adversity.
We focus on measures of exposure to violence and environmental unpredictability.
Previous research shows that these two types of adversity are on the one hand associated with improved attention shifting and working memory updating [@fields_2021; @mittal_2015; @young_2018; @young_2022], and on the other hand with lowered inhibition and working memory capacity [@fields_2021; @mittal_2015; @young_2018].
We conducted three online studies: one pilot study and two follow-up studies.
Using cognitive modeling, we unpack Flanker task performance in comparison to other tasks that require externally focused attention (Pilot study), across visual processing manipulations (Study 1), and in terms of tendencies for holistic versus local processing (Study 2).

We used an incremental preregistration approach across studies (for all preregistrations, data, code and materials, see <https://anonymous.4open.science/r/CogDev_AnonLink>).
For each study, we preregistered confirmatory (i.e., hypothesis-driven) and exploratory analyses.
The main text addresses the confirmatory analyses involving violence exposure and the exploratory analyses involving environmental unpredictability.
We describe the other exploratory analyses in the supplemental materials (section 2).
For an overview of all deviations from the preregistrations, see section 4 of the supplemental materials.

# Pilot study

In the Pilot study, our goal was to understand how childhood adversity relates to performance on tasks with different attentional demands.
Participants completed self-report measures of childhood adversity and three cognitive tasks (Flanker task, Cued Attention task, and Change Detection task).
These tasks measured inhibition, attention for peripheral cues, and attention for subtle changes.
In line with the idea that exposure to adversity may lead to a more diffuse scope of attention, we expected people with more violence exposure to be better at detecting peripheral stimuli and subtle changes.
We expected this would result in a higher drift rate (faster speed of information accumulation) or shorter non-decision times (faster attention orientation, among other things), but not necessarily with differences in boundary separation (response caution).
In contrast, we expected that participants with more violence exposure would be worse at ignoring distracting peripheral stimuli.
We expected this would result in more experienced interference (as derived from the SSP model).

@@

Participants were 565 people from the United States aged between 18 and 30 recruited on Prolific Academic ([https://www.prolific.co](https://www.prolific.co)) (See Table 1 for demographic data).
The sample was balanced on sex.
We used the MacArthur's ladder, included in Prolific's prescreening battery, for assessing perceived SES to ensure about half of the sample came from lower-SES backgrounds (which we defined as a score of 4 or below).
Participants were eligible if they spoke fluent English and did not report color-blindness.
We obtained ethical approval from the first author's institute (anonymized for peer review).

@@

We conducted a power simulation using the *faux* package in R [@debruine_2021] to determine the minimally required number of participants for standardized regression coefficients of 0.10 and 0.15 (for details and simulation code, see <https://anonymous.4open.science/r/CogDev_AnonLink/preregistrations/1_pilot/scripts>).
Power was \> .80 for adversity x task condition interactions with *N* = 450 or more.
For a linear main effect, detecting an effect of $\beta$ = 0.15 with .90 power would require *N* = `r pwr.r.test(r = 0.15, sig.level = 0.05, power = .90, alternative = "two.sided")[['n']] %>% round(0)`.
We sampled 550 participants, anticipating a final sample of \~500 after exclusions.

Prior to analyzing the data, we applied our preregistered exclusion criteria.
First, we excluded participants who did not complete the full study; second, those who had incomplete data on any of the attention tasks; third, those who missed both attention check items; fourth, those who had suspicious response patterns (e.g., consistently endorsing high response options even though some items were reverse coded).
Fifth, on a trial-level, we excluded any trials with reaction times \< 250 ms or \> 3500 ms [@ratcliff_2015].
Participants with more than 10 trials removed were completely excluded from the analyses.
The final sample consisted of `r nrow(pilot_data)` participants.

@@

The pattern of findings in the Flanker task was interesting for two reasons.
First, the Flanker task is a widely used task to assess the ability to inhibit irrelevant information, and people exposed to adversity typically show lowered performance.
Our pilot results, though, suggest that lowered performance may not be caused by a reduced ability to inhibit distracting information.
Instead, people exposed to adversity might have a lower strength of perceptual input, leading to slower and less efficient information processing.
If true, these initial findings suggest that performance might be improved through interventions that increase the visual quality of stimuli.
In Study 1, we aimed to replicate and extend these findings.

@@

The goal of Study 1 was to follow up on the Pilot study by manipulating the visual quality of information in the Flanker task.
Participants completed three versions: a standard version (similar to the Pilot study), one with enhanced visual information, and one with degraded visual information.
We again focused on childhood exposure to violence.
Our first aim was to examine the robustness of our finding of improved interference control in the Flanker task in relation to more adversity exposure in the Pilot study.
We did so by analyzing the data of the standard condition, as well as by pooling the data of the Pilot study and Study 1.
Our second aim was to investigate whether manipulating visual information in the Flanker task would influence performance for people with more violence exposure.

We preregistered two potential data patterns and associated interpretations, without favoring one over the other *a priori*.
First, the strength of perceptual input might be lower for people with more exposure to violence compared to people with less exposure to violence across all conditions.
Second, lower performance in the standard version might reflect an adaptive trade-off towards cognitive functioning that is less affected by noise or perturbations, at a cost of lower overall performance [@giudice_2018].
In that case, we would expect the strength of perceptual input to be influenced to a lesser extent across conditions for people with more exposure to violence than for people with less exposure to violence.
As a result, they might not benefit as much from enhanced visual information, yet might be able to better maintain performance with degraded information.

@@

Participant recruitment was identical to the Pilot study.
In total, 567 people from the United States between the ages of 18 and 30 participated (See Table 1).
We obtained ethical approval from the first author's institute (anonymized for peer review).
We applied the same exclusion criteria as reported in the Pilot study.
The final sample consisted of `r nrow(study1_data)` participants.

### Flanker task

We programmed the Flanker task in JsPsych version 6.3.1 [@deleeuw_2015] with three conditions.
Each condition consisted of eight practice trials, followed by 64 test trials.
In the *standard* condition, the arrows were 40 pixels in size (0.4 inches) and had zero padding between them.
In the *enhanced* condition, we increased the arrow size by 12.5% to 45 pixels (0.45 inches), and increased the space between the arrows to 5 pixels.
This increased the width of the stimulus display by 50% with respect to the standard display.
In the *degraded* condition, sizes and space between arrows were the same as in the standard version, but all arrows were rotated 45$^\circ$.
The lines of the arrows always had the same 45$^\circ$ angle.
For example, if the flanking arrows pointed to the upper-left on an incongruent trial, the central arrow pointed to the lower-right.
On congruent trials, all arrows pointed in the same direction (e.g., upper-right).
Participants completed each condition separately in different blocks, in randomized order.

@@

The self-report measures were identical to those used in the Pilot study.

@@

The procedure was identical to the Pilot study.
The full experiment took approximately 30 minutes.
Participants were paid £3.75 after they completed the full study.

@@

**Multiverse analysis.** We included the same arbitrary decisions in the multiverse analyses as in the Pilot study.
For the pooled analyses---i.e., joint analysis of the Pilot study and the standard condition of Study 1---there was one minor change in how we included screen rescaling as a preprocessing decision in the multiverse.
In Study 1, we changed the screen rescaling procedure by converting the initial size of the resize box to 300 pixels instead of 100 pixels.
This way, the stimulus display would still be close to the intended size if participants did not engage in any resizing.
However, this led to one important change for the pooled analysis: rescaling (yes or no) was included as an arbitrary exclusion decision in the multiverse analyses with four combinations: (1) exclude non-scalers in both studies; (2) include non-scalers in both studies; (3) exclude non-scalers in the Pilot study, include non-scalers in Study 1; (4) include non-scalers in the Pilot study, exclude non-scalers in Study 1.

For each analysis, we report the median $\beta$s, 95% confidence intervals, and the proportion of *p*-values \< .05 across all analytic decisions.
For the confirmatory analyses, we used bootstrapping to compute the probability of obtaining an effect size at least as extreme as observed in the real data, conditioned on a true effect size of zero [for details, see @simonsohn_2020].
See the supplemental materials (section 5) for figures summarizing *p*-distributions and the explained variance in the regression coefficients of each data cleaning decision.

**Confirmatory analyses.** To address the first aim, we analyzed the data from the standard condition, as well as pooled the Flanker task data of the Pilot study and the current study.
We ran separate linear models for each SSP parameter as well as RT difference scores (based on log-transformed mean RTs of each condition) with violence exposure as main predictor and study as covariate (effect-coded).
To address the second aim, we analyzed the effect of violence exposure and Flanker task condition type on performance using linear mixed effects models with a random intercept per participant.
The five main dependent variables were mean RT difference (based on log-transformed mean RTs of each condition) and the SSP parameters: Perceptual input, boundary separation, non-decision time, and interference.
For each outcome measure, we ran two separate models: one comparing the standard condition with the enhanced condition, and one comparing the standard condition with the degraded condition.
In both models, condition was dummy-coded using the standard condition as the reference group.

The use of RT difference scores differed from the Pilot study, where we included task condition as a moderator.
We opted for RT difference scores here (as well as in Study 2) to prevent the use of three-way interactions, for which we did not have enough power.

@@

Table 4 summarizes the multiverse results for the effects of violence exposure (confirmatory analysis) and unpredictability (exploratory analysis).
Unlike in the Pilot study, we did not find any significant associations with violence exposure.
In the exploratory analysis, there was a significant negative association between unpredictability and perceptual input (median $\beta$ = `r study1_expl_results_list$p_flanker$median_effect_Main`, 95% CI = `r study1_expl_results_list$p_flanker$CI`, `r study1_expl_results_list$p_flanker$p_sum` % of *p*s \<.05).

@@

After pooling the data of the Pilot study and Study 1, there was a negative association between violence exposure and interference (median $\beta$ = `r study1_pooled_results_list$interference_vio$median_effect`, 95% CI = `r study1_pooled_results_list$interference_vio$CI`, `r study1_pooled_results_list$interference_vio$p_sum` % of *p*s \<.05, bootstrapped *p* = `r study1_pooled_results_list$interference_vio$boot_p`).
Violence exposure was associated with lower strength of perceptual input under `r study1_pooled_results_list$interference_vio$p_sum`% of multiverse specifications, but the bootstrapped *p*-value was not significant (median $\beta$ = `r study1_pooled_results_list$p_vio$median_effect`, 95% CI = `r study1_pooled_results_list$p_vio$CI`, bootstrapped *p* = `r study1_pooled_results_list$p_vio$boot_p`).
We did not find other significant associations for either violence exposure or unpredictability.

### Flanker task conditions

@@

The results of the Pilot study and Study 1 were inconsistent with regard to the association between adversity and interference, but hinted at two general patterns.
First, violence exposure was not associated with *increased* interference; instead, we found either the opposite effect or no effect.
Second, both violence exposure and unpredictability were associated with lowered strength of perceptual input, albeit inconsistently.
These findings, if replicable, are intriguing as they would suggest that the common finding of lowered Flanker task performance among people with more adversity exposure do not actually result from worse interference control---as typically inferred.
Rather, such lowered performance would result from processes other than inhibition ability, such as slower general processing.
Though interesting, our findings so far leave open the question why adversity might be negatively associated with perceptual input.
This question was the focus of Study 2.

@@

Study 2 set out to compare two explanations for the finding that people exposed to adversity tended to show lower strength of perceptual input in the Flanker task.
First, lowered strength of perceptual input in people exposed to adversity may indicate a difficulty in extracting relevant information (i.e., about their direction) from the arrows in general.
Second, the difference in perceptual input may not be a cognitive deficit per se, but instead could be a signature of a difference in processing style---that is, a feature, and not a bug.
People exposed to adversity may process information more holistically, focusing more on the configuration of pieces of information rather than individual pieces of information.
in the Flanker task, this would lower the depth of perceptual processing of any individual stimulus, thus resulting in lowered strength of perceptual input, as we observed in the Pilot study and Study 1.

We preregistered three aims focusing both on violence exposure and unpredictability.
First, we expected to replicate our earlier findings that adversity was associated with lowered perceptual input and lower interference in people exposed to adversity.
Second, we included a Global-Local task to investigate the hypothesis---based on the findings of the Pilot study and Study 1---that people with more adverse experiences would develop a more holistic style of information processing.
Third, we planned to conduct a within-subjects analysis of Flanker and Global-Local task performance to assess whether people with lowered perceptual input in the Flanker task would also show a more global processing style (rather than a local processing style) in the Global-Local task.

@@

Participants were 600 people from the United States between the ages of 18 and 30.
Recruitment was identical to Study 1.
We obtained ethical approval from the first author's institute (anonymized for peer review).
We conducted a simulation-based power analysis for the planned linear mixed models with the Global Local task (see GitHub).
We determined that power of \> .80 for a standardized interaction effect of 0.06, with sigma (noise) set to 0.7 (comparable to observed sigmas in the first two studies) would require 550 participants.
We recruited 600 participants, with the expectation to have a final sample of 550 participants after exclusions.
We applied the same exclusion criteria as reported in the Pilot study and Study 1.
The final sample consisted of `r nrow(study2_data)` participants.

@@

The measures of childhood violence exposure and environmental unpredictability were identical to Study 1.
The Flanker task was identical to the standard version used in Study 1.

*Global-Local task*.
The Global-Local task is a measure of global-local processing [@navon_1977].
Many different versions of this task exist in the literature.
One key dimension on which they differ is whether the task measures focused attention (by cueing attention towards the global or local level prior to stimulus presentation) or divided attention (by having participants search for a target on both levels) [@lee_2023].
Here, we use a version measuring divided attention, which allows measuring whether someone tends to have a more global versus local processing style [@hakim_2017; @lee_2023; @mckone_2010].

Participants saw images of big, black letters (the global level) comprising small letters (the local level)---so-called Navon images [@navon_1977]---against a white background.
Participants first completed eight practice trials, after which they completed an additional 64 test trials.
On each trial, participants searched for one of two target letters---an 'E' or 'H'---and indicated whether it was present on the global or local level by pressing 'g' or 'l' on their keyboard, respectively.
Each stimulus was 600 pixels high and 395 pixels wide and comprised seven local letters vertically and five local letters horizontally.
The stimuli consisted of combinations of the letters 'T', 'F', 'P', 'L', 'H', and 'E'.
All stimuli always contained one (and only one) of the target letters 'H' and 'E' on either the local or global level.
The other letters were randomly varied, and the global and local level never contained the same letter.
Thus, the global-local task did not contain a congruent and incongruent condition as did the Flanker task.

@@

**Multiverse analysis.** We included the same decisions in the multiverse as in the previous studies.
However, there was one deviation from the preregistration: the multiverse analysis contained the same arbitrary decisions as the Pilot study and Study 1, instead of a subset, as we preregistered (for details, <https://anonymous.4open.science/r/CogDev_AnonLink/preregistrations/3_study2>).
See the supplemental materials (section 5) for figures summarizing *p*-distributions and the explained variance in the regression coefficients of each data cleaning decision.

**DDM estimation.** For the Flanker task, we used the SSP [@grange_2016; @white_2018a; @white_2018b; @white_2011] using the same procedure as in Study 1. For the Global-Local task, we used a hierarchical Bayesian DDM to fit the data using the *runjags* package [@denwood_2016]. See the supplemental materials (Section 3) for more information about the procedure and model fit.

We deviated from our preregistration regarding the preprocessing of Global-Local task data.
Specifically, we relaxed the low performance threshold as the task was more difficult than anticipated.
These deviations are described in the supplemental materials (section 4).

**Confirmatory analyses.** We ran simple regressions for analyses involving only main effects (aim 1), and linear mixed effects models for analyses involving within-subject interactions (aim 2 and 3).
To address aim 3 (within-subject interaction between Global-Local task drift rate and Flanker task strength of perceptual input), we further preprocessed the data in two steps.
First, we computed a difference score of Global-Local drift rates by subtracting the drift rate on local trials from the drift rate on global trials (with higher scores reflecting relatively faster information processing on global trials).
Second, we separately standardized the Flanker task strength of perceptual input and Global-Local task drift rate difference.
We fitted linear mixed effects models with the standardized performance measures as the dependent variable, and adversity type, task (Flanker task or Global-Local task, sum-coded) and their interaction as independent variables.

@@

```{r}
#| label: figure3
#| fig.width: 7
#| fig.height: 9
#| dpi: 600
#| fig-cap: | 
#|   **Figure 3.** Multiverse results for the association between unpredictability with the strength of perceptual input and interference in the Flanker task across all studies. Each panel depicts sorted beta coefficients across all combinations of arbitrary decisions (i.e., the effect curve across the whole multiverse). The top row depicts effect curves in the Pilot study. The second row depicts effect curves in Study 1. The third row depicts effect curves in Study 2. The Fourth row depicts effect curves of the pooled analyses across all studies.
knitr::include_graphics("figures/fig3.png")
```

@@

### Global-Local task performance

@@

To sum up, Study 2 provided additional support for the basic finding that violence exposure and unpredictability were associated with lower strength of perceptual input but not with differences in interference; with the caveat that the associations for unpredictability only showed up in pooled analyses.
People with more exposure to violence and unpredictability also processed information more slowly in the Global-Local task.
In line with our expectations, childhood exposure to violence was associated with both lowered strength of perceptual input and a more holistic processing style.
The same processing style was observed for participants with more exposure to unpredictability, although they did not show lowered strength of perceptual input.

@@

**Drift Diffusion Model:** A formal cognitive model that explicitly models the cognitive processes underlying decision-making on speeded binary decision-making tasks.
It represents decision-making as a process in which people accumulate information until one response is sufficiently favored over the other.

@@

**Perceptual input:** Parameter of the Shrinking Spotlight Model of Flanker task performance. Measures how well a person can process the Flanker task stimuli in general.

@@

**Shrinking Spotlight Model:** An extension of the standard Drift Diffusion Model to account for attention processes in the Flanker task.
